---
title: "Olympics medalists, not by country, but by NUTS region"
author:
  - name: Giorgio Comai
    url: https://giorgiocomai.eu
    affiliation: OBCT/CCI - EDJNet
    affiliation_url: https://europeandatajournalism.eu/
    orcid_id: 0000-0002-0515-9542
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    includes:      
      in_header: header.html
---


```{r load_packages, messages = FALSE, warning = FALSE, echo = FALSE}
source(here::here("functions", "setup.R"))
source(here::here("functions", "extract_medals_from_wikipedia_page.R"))
source(here::here("functions", "get_list_of_lists.R"))
```

## What this is about

The Olympics medals table is notoriously all about countries and national teams. But what if the Olympic medal table was based on the number of medals won by regions, not by countries?

We check this based on the place of birth for all athletes born in one of the European NUTS regions; [NUTS](https://en.wikipedia.org/wiki/Nomenclature_of_Territorial_Units_for_Statistics) - Nomenclature of Territorial Units for Statistics - is a standardised classification of administrative entities defined by the European Union ([here’s the geographic dataset for download](https://ec.europa.eu/eurostat/web/gisco/geodata/statistical-units/territorial-units-statistics)).

We conduct this analysis for the 2024 Summer Olympics, and we checked these data for the 2020 Summer Olympics. But in principle the procedure can be expanded in order to include *all* modern Olympics, all over the world. 

A map including all known places of birth including also non-European-born athletes, the corresponding dataset, and all code used to generate them are [shared in this repository](https://github.com/EDJNet/OlympicsGoNUTS).

## Haven't I seen this before?

This is an updated version of a similar excercise I conducted for the 2020 Summer Olympics. 

- [Ranking European regions by Olympics medals](https://www.europeandatajournalism.eu/cp_data_news/ranking-european-regions-by-olympics-medals/)
- [The data you need to win the Olympics if you go NUTS](https://www.europeandatajournalism.eu/cp_data_news/the-data-you-need-to-win-the-olympics-if-you-go-nuts/)
- [repository with data and code](https://github.com/EDJNet/olympics2020nuts) / [web version](https://edjnet.github.io/olympics2020nuts/)
- [interactive full screeen map](https://edjnet.github.io/olympics2020nuts/medalists_map.html) with place of birth of all medalists at the 2020 Summer Olympics

## How do we go about it?

This whole data retrieval process is based on Wikipedia and Wikidata. Wikidata provides data in a more machine-readable format, but, at least in the past, the Wikipedia page listing all medalists was more complete and was updated much more quickly (there's a [WikiProject Olympics](https://en.wikipedia.org/wiki/Wikipedia:WikiProject_Olympics), if you're looking for inputs on how to contribute). In brief, we get medalists from the Wikipedia page, then retrieve data about their place of birth from Wikidata, then with some geo-matching we attribute each of them to a NUTS region.

We can then combine these data with other statistics about these regions, such as population and economic indicators.

## Datasets and maps

**Warning**: Provisional data, not checked for accuracy!

```{r link_to_all_html, results='asis', echo=FALSE}
years_v <- list_of_lists_df |> 
  dplyr::filter(year>1945) |> 
  dplyr::pull(year)

purrr::walk(years_v, \(x) {
  cat(stringr::str_c("- [", x, " Summer Olympics]","(", x,")"))
  cat("\n")
})
```


## All medalists on Wikipedia

Wikipedia has a [category page](https://en.wikipedia.org/wiki/Category:Lists_of_Summer_Olympic_medalists_by_year) that includes links to all Wikipedia pages that report all Olympics medal winners for all modern summer Olympics. Here they are:

```{r list_of_lists, echo=FALSE}


list_of_lists_df |> 
  dplyr::transmute(year,
                   title = kableExtra::cell_spec(title_url, link = url)
                   # ,
                   # qid = kableExtra::cell_spec(qid, link = stringr::str_c("https://www.wikidata.org/wiki/", qid))
                   ) |> 
  dplyr::arrange(as.numeric(year)) |> 
  knitr::kable(escape = FALSE)
```


## Credits

This dataset has been generated by Giorgio Comai (OBCT/CCI) within the scope of EDJNet, the European Data Journalism Network. It is released under a CC-BY license (Giorgio Comai/OBCT/EDJNet). Please do include a reference to EDJNet - the European Data Journalism Network if you use these data.
